pipeComp, a general framework for the evaluation of computational pipelines, reveals performant single cell RNA-seq preprocessing tools

### 2020-05-06 bmc_article.tex

Here, we develop a flexible R framework for pipeline comparison and evaluate the various steps of analysis leading from an initial count matrix to a cluster assignment, which are critical in a wide range of applications. We collected real datasets of known cell composition (Table \ref{tab:table1}) and used a variety of evaluation metrics to investigate in a multilevel fashion the impact of various parameters and variations around a core scRNAseq pipeline. Although we use some datasets based on other protocols, our focus is especially on droplet-based datasets that do not include exogenous control RNA (i.e. spike-ins); see Table \ref{tab:table1} and Figure \ref{fig:figure1} for more details. In addition to previously-used benchmark datasets with true cell labels \cite{duoClustering2018,tianMixology2018}, we simulated two datasets with a hierarchical subpopulation structure based on real 10x human and mouse data using \textit{muscat} \cite{CrowellMuscat2019}. 
Since graph-based clustering \cite{satijaSeurat2015} was previously shown to consistently perform well across several datasets \cite{duoClustering2018,tianMixology2018}, we used the \textit{Seurat} pipeline as the starting point to perform an integrated investigation of: 1) doublet identification; 2) cell filtering; 3) normalization; 4) feature selection; 5) dimension reduction; and, 6) clustering. We compared competing approaches and also explored more fine-grained parameters and variations on common methods. Importantly, the success of methods at a certain analytical step might be dependent on choices at other steps. Therefore, instead of evaluating each step in isolation, we developed a general framework for evaluating nested variations on a pipeline and suggest a multilevel panel of metrics. Finally, we evaluate several recent methods and provide practical recommendations.

@@

The \textit{pipeComp} package defines a pipeline as, minimally, a list of functions executed consecutively on the output of the previous one (Figure \ref{fig:figure2}A). In addition, optional benchmark functions can be set for each step to provide standardized, multi-layered evaluation metrics. Given such a \texttt{PipelineDefinition} object, a set of alternative parameters (which might include different subroutines) and benchmark datasets, the \texttt{runPipeline} function then proceeds through all combinations of arguments, avoiding recomputing the same step twice and compiling evaluations (including run time) on the fly. Variations in a given parameter can be evaluated using all metrics from this point downward in the pipeline. This is especially important because end-point metrics, such as the adjusted Rand index (ARI) \cite{HubertARI1985} for clustering, are not perfect. For example, although the meaning of an ARI score is independent of the number of true subpopulations \cite{steinleyProperties2004}, the number of clusters called is by far the most important determinant of the score: the farther it is from the actual number of subpopulations, the worse the ARI (Supplementary Figures 1-2). In this context, one strategy has been to cluster across various resolutions and only consider the results that have the right number of clusters \cite{duoClustering2018}. While this has the virtue of making the results comparable, in practice the number of subpopulations is typically unknown and tools that operate well in this optimal context might not necessarily be best overall. Clustering results are also very sensitive and might not always capture improvements in earlier steps of the pipeline. In addition, we wanted to assess whether the effect of a given parameter alteration is robust to changes in the rest of the pipeline. We therefore monitored several complementary metrics across multiple steps of the process. We proceeded in a step-wise fashion, first testing a large variety of parameters at the early steps of the pipeline along with only a set of mainstream options downstream, then selecting main alternatives and proceeding to a more detailed benchmark of the next step (Figure \ref{fig:figure2}B).

@@

@@

On the basis of our findings, we can make a number of practical analysis recommendations, also summarized in Figure \ref{fig:figure11}:

@@

\textit{pipeComp} is a flexible R framework for evaluating methods and parameter combinations in a complex pipeline, computing on-the-fly multilevel evaluation metrics. Applying this framework to scRNAseq clustering enabled us to make clear recommendations on the steps of filtering, normalization, feature selection, denoising, dimensionality reduction and clustering (Figure \ref{fig:figure11}). We demonstrate how a diversity of multilevel metrics can be more robust, more sensitive, and more nuanced than simply evaluating the final clustering performance. We hope that the \textit{pipeComp} framework can be applied to extend the current field of benchmarking, as well as to investigate pipelines other domains with multi-step computational analyses.

@@

The \textit{stringent} set of filters uses the same thresholds, but excludes a cell if it is an outlier on any single distribution.

The \textit{lenient} set of filters excludes cells that are outliers on at least two distributions by at least 5 MADs, except for pct\_counts\_Mt where the threshold is $>$ 3 MADs and $>$ 0.08.

@@

@@

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[scale=0.75, keepaspectratio]{main_figures/pipeline_explanation.pdf}
    \fi
    \caption{\textbf{Overview of the \textit{pipeComp} framework and its application to a scRNAseq clustering pipeline. A:} The package is built around a \texttt{PipelineDefinition} S4 class which defines a set of functions to be executed consecutively, as well as optional evaluation functions for each step. Each step can accept a number of parameters whose alternative values are provided as a list. All (or subsets of) combinations of parameters can then be simultaneously ran and evaluated using the \texttt{runPipeline} function. \textbf{B:} Scheme representing the application of \textit{pipeComp} to evaluate a range of methods that are commonly used in scRNAseq studies and some of the metrics monitored at various steps.}
    \label{fig:figure2}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[scale=0.85, keepaspectratio]{{main_figures/doublets.pdf}}
    \fi
    \caption{\textbf{Identification of doublet cells. A:} Receiver operating characteristic (ROC) curves of the tested doublet detection methods for three datasets with SNP-identified doublets. Dots indicate threshold determined by the true number of doublets. \textbf{B:} Running time of the different methods (\textit{DoubletFinder} failed on one of the datasets). \textbf{C:} Rate of misclassification of the cells identified by \textit{scDblFinder} as doublets (DBL) or singlets (SNG), across a large range of clustering analysis. Even in datasets which should not have neotypic doublets, the cells identified as such tend to be misclassified.    \label{fig:figure3}}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[scale=0.72, keepaspectratio]{{main_figures/filtering.pdf}}
    \fi
    \caption{\textbf{Effect of filtering on cell subpopulation structure and clustering. A:} Filtering on the basis of distance to the whole distribution can lead to strong bias against certain subpopulations. The dashed line indicates a threshold of 2.5 median absolute deviations (MADs) from the median of the overall population. \textbf{B:} Relationship between the maximum subpopulation exclusion rate and the average clustering accuracy per subpopulation across various filtering strategies. Of note, doublet removal appears to be desirable even when, due to the design, there are no heterotypic doublets in the data. The PCA methods refer to multivariate outlier detected as implemented in \textit{scater} (see Methods for details).}
    \label{fig:figure4}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[scale=0.6, keepaspectratio]{{main_figures/normalization.pdf}}
    \fi
    \caption{\textbf{Evaluation of normalization procedures. A:} Variance in PC1 explained by library size (left) and detection rate (right) after accounting for subpopulation differences. \textbf{B:} Average silhouette width per true subpopulation, where higher silhouette width means a higher separability. \textbf{C:} Clustering accuracy (\textit{Seurat} clustering), measured by mutual information (MI), Adjusted Rand Index (ARI) and ARI at the true number of cluster. Grey squares indicate that the true number of clusters could not be reached with the corresponding method and dataset. `feat\_regress', `mt\_regress' and `feat\_mt\_regress' respectively stand for the regressing out of the number of features, the proportion of mitochondrial reads, or both during scaling.}
    \label{fig:figure5}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[scale=0.65]{{main_figures/selection.pdf}}
    \fi
    \caption{\textbf{Evaluation of feature selection methods. A:} Ability of different feature ranking methods to capture genes with a high proportion of variance (left) or deviance (right) explained by real subpopulations. The asterisk denotes the default Seurat method. \textbf{B:} Accuracy of clusterings (at the true number of clusters) when selecting 1000 genes using the given methods. Based on standard \textit{Seurat} normalization (left) or \textit{SCTransform} (right). The  \textit{vst.varExp} and \textit{devianceExplained} methods correspond to the estimates used in \textbf{A} to evaluate the selection methods and were included here only for validation purpose.}
    \label{fig:figure6}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[scale=0.50, keepaspectratio]{{main_figures/dr.pdf}}
    \fi
    \caption{\textbf{Evaluation of common dimensionality reduction methods. A:} Average silhouette width per subpopulation resulting from combinations of normalization and dimension reductions. \textbf{B:} Clustering accuracy, mean proportion of the variance in the first 5 components explained by real subpopulations (center), and median ARI of the resulting clustering (right).}
    \label{fig:figure7}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[scale=0.53, keepaspectratio]{{main_figures/dimensionality.pdf}}
    \fi
    \caption{\textbf{Estimating dimensionality. A:} Estimated dimensionality using the proportion of variance in each component explained by the true subpopulations. \textbf{B:} Difference between `real' dimensionality (from \textbf{A}) and dimensionality estimation methods, along with computing time. \textbf{C:} Average silhouette width per subpopulation across a selection of methods, combined with \textit{SCTransform} or standard \textit{Seurat} normalization. \textbf{D:} Clustering accuracy across the normalization/dimensionality estimation methods.}
    \label{fig:figure8}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[scale=0.43, keepaspectratio]{{main_figures/clustering.pdf}}
    \fi
    \caption{\textbf{Evaluation of clustering methods.} Clustering accuracy (\textbf{A}) and run time (\textbf{B}) of \textit{scran}- and \textit{Seurat}-based clustering tools in combination with \textit{SCTransform} and standard \textit{Seurat}'s normalization.}
    \label{fig:figure9}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[scale=0.55, keepaspectratio]{{main_figures/imputation.pdf}}
    \fi
    \caption{\textbf{Evaluation of imputation/denoising methods} Average silhouette width per subpopulation (\textbf{A}) and clustering accuracy (\textbf{B}) with or without (indicated as \textit{none}) application of a denoising/imputation method.}
    \label{fig:figure10}
\end{figure}

\begin{figure}[h]
    \centering
    \ifshowfigs
    \includegraphics[scale=0.9,keepaspectratio]{main_figures/pipeComp_summary.pdf}
    \fi
    \caption{\textbf{Recommendations of tools from filtering to clustering.}}
    \label{fig:figure11}
\end{figure}

@@

\begin{table}[h!]
\caption{Overview of the benchmark datasets.}
\label{tab:table1}
\begin{tabular}{rlrrl}
  \hline
dataset & source & # cells & protocol & description \\ 
  \hline
Koh & GSE85066 & 531 & SMARTer & FACS purified differentiation stages \\ 
  Kumar & GSE60749 & 246 & SMARTer & Mouse ESC cultured in different conditions \\ 
  Zhengmix4eq & \cite{duoClustering2018} & 3994 & 10x & Mixtures of FACS purified PBMCs \\ 
  Zhengmix4uneq & \cite{duoClustering2018} & 6498 & 10x & Mixtures of FACS purified PBMCs \\ 
  Zhengmix8eq & \cite{duoClustering2018} & 3994 & 10x & Mixtures of FACS purified PBMCs \\ 
  mixology10x3cl & \cite{tianMixology2018} & 902 & 10x & Mixture of 3  cancer cell lines from CellBench \\ 
  mixology10x5cl & \cite{tianMixology2018} & 3918 & 10x & Mixture of 5 cancer cell lines from CellBench \\ 
  simMix1 & - & 2500 & 10x-based & Simulation of 10 human cell subpopulations \\
  simMix2 & - & 3000 & 10x-based & Simulation of 9 mouse cell subpopulations \\
   \hline
\end{tabular}
\end{table}

>>>>

Here, we develop a flexible R framework for pipeline comparison and apply it to the evalution of the various steps of analysis leading from an initial count matrix to a cluster assignment, which are critical in a wide range of applications. We collected real datasets of known cell composition (Table \ref{tab:table1}) and used a variety of evaluation metrics to investigate in a multilevel fashion the impact of various parameters and variations around a core scRNAseq pipeline. Although we use some datasets based on other protocols, our focus is especially on droplet-based datasets that do not include exogenous control RNA (i.e. spike-ins); see Table \ref{tab:table1} and Figure \ref{fig:figure1} for more details. In addition to previously-used benchmark datasets with true cell labels \cite{duoClustering2018,tianMixology2018}, we simulated two datasets with a hierarchical subpopulation structure based on real 10x human and mouse data using \textit{muscat} \cite{CrowellMuscat2019}. 
Since graph-based clustering \cite{satijaSeurat2015} was previously shown to consistently perform well across several datasets \cite{duoClustering2018,tianMixology2018}, we used the \textit{Seurat} pipeline as the starting point to perform an integrated investigation of: 1) doublet identification; 2) cell filtering; 3) normalization; 4) feature selection; 5) dimension reduction; and, 6) clustering. We compared competing approaches and also explored more fine-grained parameters and variations on common methods. Importantly, the success of methods at a certain analytical step might be dependent on choices at other steps. Therefore, instead of evaluating each step in isolation, we developed a general framework for evaluating nested variations on a pipeline and suggest a multilevel panel of metrics. Finally, we evaluate several recent methods and provide practical recommendations.

@@

The \textit{pipeComp} package defines a pipeline as, minimally, a list of functions executed consecutively on the output of the previous one (Figure \ref{fig:figure2}A). In addition, optional benchmark functions can be set for each step to provide standardized, multi-layered evaluation metrics. Given such a \texttt{PipelineDefinition} object, a set of alternative parameters (which might include different subroutines) and benchmark datasets, the \texttt{runPipeline} function then proceeds through all combinations of arguments (or a desired subset), avoiding recomputing the same step twice and compiling evaluations (including run time) on the fly (see Supplementary methods for more detail). Variations in a given parameter can be evaluated using all metrics from this point downward in the pipeline. This is especially important because end-point metrics, such as the adjusted Rand index (ARI) \cite{HubertARI1985} for clustering, are not perfect. For example, although the meaning of an ARI score is independent of the number of true subpopulations \cite{steinleyProperties2004}, the number of clusters called is by far the most important determinant of the score: the farther it is from the actual number of subpopulations, the worse the ARI (Supplementary Figures 1-2). In this context, one strategy has been to cluster across various resolutions and only consider the results that have the right number of clusters \cite{duoClustering2018}. While this has the virtue of making the results comparable, in practice the number of subpopulations is typically unknown and tools that operate well in this optimal context might not necessarily be best overall. Clustering results are also very sensitive and might not always capture improvements in earlier steps of the pipeline. In addition, we wanted to assess whether the effect of a given parameter alteration is robust to changes in the rest of the pipeline. We therefore monitored several complementary metrics across multiple steps of the process. We proceeded in a step-wise fashion, first testing a large variety of parameters at the early steps of the pipeline along with only a set of mainstream options downstream, then selecting main alternatives and proceeding to a more detailed benchmark of the next step (Figure \ref{fig:figure2}B).

@@

\subsection*{Application of _pipeComp_ to a different context}

To illustrate the applicability of \textit{pipeComp} to a completely different setting, we benchmarked the impact of surrogate variable analysis (SVA) -based methods on bulk RNAseq differential expression analysis (DEA). The rationale behind such approaches is that the identification of factors determining genome-wide variation (i.e. surrogate variables) and their inclusion in the generalized linear model for DEA might improve DEA, especially in the presence of sources of variation unrelated to the experimental variables. \textit{sva}\citep{leekCapturingHeterogeneityGene2007,leekSvaPackageRemoving2012} and \textit{RUVseq}\citep{rissoNormalizationRNAseq2014a} are two popular packages implementing methods to infer these variables. We therefore created a \texttt{PipelineDefinition} with three step: 1) feature filtering, 2) removal of unwanted variation (i.e. SVA), and 3) DEA using the most established methods (the package's `pipeComp\_dea' vignette details the creation process). We developed 3 benchmark datasets: 1) the \texttt{ipsc} data is a random selection of 10 vs 10 samples from \citet{carcamoOriveAnalysisTranscriptional2017}, which are very heterogeneous, and foldchanges were manually added to 300 genes; 2) the \texttt{seqc} dataset contains 5 vs 5 samples of two mixtures of the SEQC consortium with different ERCC spike-in mixes, and includes a partially correlated batch effect; 3) A **simulation** of 8 vs 8 samples with two technical vectors of variation (see methods for more details). As previously reported, we found that more stringent filtering improved DEA (Supplementary Figure 25A) and that the different DEA methods had similar performances (Figure \ref{fig:svadea}A and Supplementary Figure 25A). Including a SVA-based step increased sensitivity across all datasets, and particularly so in datasets with technical vectors of variation, and benefited all DEA methods in a similar fashion (Figure \ref{fig:svadea}A,C and Supplementary Figures 25-26). With the exception of the \texttt{RUVr} method, which led to a high False Discovery Rate (FDR), all other methods showed comparable gains in sensitivity while maintaining a decent error control (if a little above the nominal FDR), even when increasing the number of surrogate variables used (Figure \ref{fig:svadea}B-C). These results therefore suggest that these methods are robust and can be widely applied.

@@

On the basis of our findings, we can make a number of practical analysis recommendations, also summarized in Figure \ref{fig:recommendations}:

@@

\textit{pipeComp} is a flexible R framework for evaluating methods and parameter combinations in a complex pipeline, computing on-the-fly multilevel evaluation metrics. Applying this framework to scRNAseq clustering enabled us to make clear recommendations on the steps of filtering, normalization, feature selection, denoising, dimensionality reduction and clustering (Figure \ref{fig:recommendations}). We demonstrate how a diversity of multilevel metrics can be more robust, more sensitive, and more nuanced than simply evaluating the final clustering performance. We hope that the \textit{pipeComp} framework can be applied to extend the current field of benchmarking, as well as to investigate pipelines other domains with multi-step computational analyses.

@@

The \textit{stringent} set of filters uses the same thresholds, but excludes a cell if it is an outlier on any single distribution. 
The \textit{veryStringent} set of filters excludes any cell that is an outlier by at least 2 MADs (lower or higher) on any of: log10\_total\_counts, log10\_total\_features, pct\_counts\_Mt, or pct\_counts\_in\_top\_20\_features.
The \textit{lenient} set of filters excludes cells that are outliers on at least two distributions by at least 5 MADs, except for pct\_counts\_Mt where the threshold is $>$ 3 MADs and $>$ 0.08.

@@

\subsection*{SVA-DEA benchmark}

@@

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[width=\textwidth, keepaspectratio]{main_figures/pipeline_explanation.pdf}
    \fi
    \caption{\textbf{Overview of the \textit{pipeComp} framework and its application to a scRNAseq clustering pipeline. A:} The package is built around a \texttt{PipelineDefinition} class which defines a set of functions to be executed consecutively, as well as optional evaluation functions for each step. Each step can accept a number of parameters whose alternative values are provided as a list. All (or subsets of) combinations of parameters can then be simultaneously ran and evaluated using the \texttt{runPipeline} function (see Supplementary Methods for more details). \textbf{B:} Scheme representing the application of \textit{pipeComp} to evaluate a range of methods that are commonly used in scRNAseq studies and some of the metrics monitored at various steps.}
    \label{fig:figure2}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[width=\textwidth, keepaspectratio]{{main_figures/doublets.pdf}}
    \fi
    \caption{\textbf{Identification of doublet cells. A:} Receiver operating characteristic (ROC) curves of the tested doublet detection methods for three datasets with SNP-identified doublets. Dots indicate threshold determined by the true number of doublets. \textbf{B:} Running time of the different methods (\textit{DoubletFinder} failed on one of the datasets). \textbf{C:} Rate of misclassification of the cells identified by \textit{scDblFinder} as doublets (DBL) or singlets (SNG), across a large range of clustering analysis. Even in datasets which should not have neotypic doublets, the cells identified as such tend to be misclassified.    \label{fig:figure3}}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[width=\textwidth, keepaspectratio]{{main_figures/filtering.pdf}}
    \fi
    \caption{\textbf{Effect of filtering on cell subpopulation structure and clustering. A:} Filtering on the basis of distance to the whole distribution can lead to strong bias against certain subpopulations. The dashed line indicates a threshold of 2.5 median absolute deviations (MADs) from the median of the overall population. \textbf{B:} Relationship between the maximum subpopulation exclusion rate and the average clustering accuracy per subpopulation across various filtering strategies. Of note, doublet removal appears to be desirable even when, due to the design, there are no heterotypic doublets in the data. The PCA methods refer to multivariate outlier detected as implemented in \textit{scater} (see Methods for details).}
    \label{fig:figure4}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[width=\textwidth, keepaspectratio]{{main_figures/normalization.pdf}}
    \fi
    \caption{\textbf{Evaluation of normalization procedures. A:} Variance in PC1 explained by library size (left) and detection rate (right) after accounting for subpopulation differences. \textbf{B:} Average silhouette width per true subpopulation, where higher silhouette width means a higher separability. \textbf{C:} Clustering accuracy (\textit{Seurat} clustering), measured by mutual information (MI), Adjusted Rand Index (ARI) and ARI at the true number of cluster. Grey squares indicate that the true number of clusters could not be reached with the corresponding method and dataset. `feat\_regress', `mt\_regress' and `feat\_mt\_regress' respectively stand for the regressing out of the number of features, the proportion of mitochondrial reads, or both during scaling.}
    \label{fig:figure5}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[width=\textwidth]{{main_figures/selection.pdf}}
    \fi
    \caption{\textbf{Evaluation of feature selection methods. A:} Ability of different feature ranking methods to capture genes with a high proportion of variance (left) or deviance (right) explained by real subpopulations. The asterisk denotes the default Seurat method. \textbf{B:} Accuracy of clusterings (at the true number of clusters) when selecting 1000 genes using the given methods. Based on standard \textit{Seurat} normalization (left) or \textit{SCTransform} (right). The  \textit{vst.varExp} and \textit{devianceExplained} methods correspond to the estimates used in \textbf{A} to evaluate the selection methods and were included here only for validation purpose.}
    \label{fig:figure6}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[width=\textwidth, keepaspectratio]{{main_figures/dr.pdf}}
    \fi
    \caption{\textbf{Evaluation of common dimensionality reduction methods. A:} Average silhouette width per subpopulation resulting from combinations of normalization and dimension reductions. \textbf{B:} Clustering accuracy, mean proportion of the variance in the first 5 components explained by real subpopulations (center), and median ARI of the resulting clustering (right).}
    \label{fig:figure7}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[width=\textwidth, keepaspectratio]{{main_figures/dimensionality.pdf}}
    \fi
    \caption{\textbf{Estimating dimensionality. A:} Estimated dimensionality using the proportion of variance in each component explained by the true subpopulations. \textbf{B:} Difference between `real' dimensionality (from \textbf{A}) and dimensionality estimation methods, along with computing time. \textbf{C:} Average silhouette width per subpopulation across a selection of methods, combined with \textit{SCTransform} or standard \textit{Seurat} normalization. \textbf{D:} Clustering accuracy across the normalization/dimensionality estimation methods.}
    \label{fig:figure8}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[width=\textwidth, keepaspectratio]{{main_figures/clustering.pdf}}
    \fi
    \caption{\textbf{Evaluation of clustering methods.} Clustering accuracy (\textbf{A}) and run time (\textbf{B}) of \textit{scran}- and \textit{Seurat}-based clustering tools in combination with \textit{SCTransform} and standard \textit{Seurat}'s normalization.}
    \label{fig:figure9}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[width=\textwidth, keepaspectratio]{{main_figures/imputation.pdf}}
    \fi
    \caption{\textbf{Evaluation of imputation/denoising methods} Average silhouette width per subpopulation (\textbf{A}) and clustering accuracy (\textbf{B}) with or without (indicated as \textit{none}) application of a denoising/imputation method.}
    \label{fig:figure10}
\end{figure}

\begin{figure}
    \centering
    \ifshowfigs
    \includegraphics[width=\textwidth, keepaspectratio]{{main_figures/svadea.pdf}}
    \fi
    \caption{\textbf{Evaluation the impact of prior factor analysis on (bulk) RNAseq differential expression analysis using \textit{pipeComp}. A:} True positive rates (TPR) and False discovery rates (FDR) of various combinations of SVA-based methods (using a single surrogate variable) and DEA methods. \textbf{B:} TPR and FDR of the different SVA-based methods (averaged across the different DEA methods) using increasing numbers of variables (`be' and `leek` refer to the two methods implemented in the \textit{sva} package for estimating the number of surrogate variables). In \textbf{A-B}, the x-axis is square-root transformed to improve readability, and three dots for each combination of methods refer to the nominal FDR thresholds of 0.01, 0.05, and 0.1. \textbf{C:} Correlation of estimated logFC with expected ones, as well as TPR and FDR (at nominal FDR threshold of 0.05) of the different methods with increasing numbers of variables.}
    \label{fig:svadea}
\end{figure}

\begin{figure}[h]
    \centering
    \ifshowfigs
    \includegraphics[scale=0.9,keepaspectratio]{main_figures/pipeComp_summary.pdf}
    \fi
    \caption{\textbf{Recommendations of tools from filtering to clustering.}}
    \label{fig:recommendations}
\end{figure}

@@

\begin{table}[h!]
\caption{Overview of the benchmark datasets.}
\label{tab:table1}
\begin{tabular}{rlrrl}
  \hline
dataset & source & \# cells & protocol & description \\ 
  \hline
Koh & GSE85066 & 531 & SMARTer & FACS purified differentiation stages \\ 
  Kumar & GSE60749 & 246 & SMARTer & Mouse ESC cultured in different conditions \\ 
  Zhengmix4eq & \cite{duoClustering2018} & 3994 & 10x & Mixtures of FACS purified PBMCs \\ 
  Zhengmix4uneq & \cite{duoClustering2018} & 6498 & 10x & Mixtures of FACS purified PBMCs \\ 
  Zhengmix8eq & \cite{duoClustering2018} & 3994 & 10x & Mixtures of FACS purified PBMCs \\ 
  mixology10x3cl & \cite{tianMixology2018} & 902 & 10x & Mixture of 3  cancer cell lines from CellBench \\ 
  mixology10x5cl & \cite{tianMixology2018} & 3918 & 10x & Mixture of 5 cancer cell lines from CellBench \\ 
  simMix1 & - & 2500 & 10x-based & Simulation of 10 human cell subpopulations \\
  simMix2 & - & 3000 & 10x-based & Simulation of 9 mouse cell subpopulations \\
   \hline
\end{tabular}
\end{table}
